---
title: "DW Ex1 Basic Data Manipulation"
author: "Barry Davis"
date: "June 21, 2017"
output: 
  html_document:
    code_folding: none
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Set up our environment
I'll be using some functions from dplyr and tidyr, so I will install the tidyverse package. This package includes dplyr and tidyr, along with other packages that can/may be helpful. 

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
```

## Load the data into RStudio
Next, I load the data from the external Excel file that was provided into a data frame.
```{r}
prodPurchases <- read.csv("refine_original.csv", header = TRUE, sep = ",")
```
Let's take a peak and see what's inside.
```{r echo=FALSE}
kable(head(prodPurchases)) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```

## Time to clean it up
The names of the companies in this file were misspelled and had irregular capitalization. I will start by correcting those fields first. Regex can be a fickle beast, so I will first verify, then change.
```{r}
# Change to lowercase
prodPurchases$company <- tolower(prodPurchases$company)

# Test the regex, then change what matched for company names
test <- grep(pattern = ".*\\ps$", prodPurchases$company)
for(value in test) {prodPurchases[value,1] <- sub(pattern = ".*\\ps$", replacement = "Philips", x = prodPurchases[value,1])}

test <- grep(pattern = "^ak.*", prodPurchases$company)
for(value in test) {prodPurchases[value,1] <- sub(pattern = "^ak.*", replacement = "Akzo", x = prodPurchases[value,1])}

test <- grep(pattern = "^u.*", prodPurchases$company)
for(value in test) {prodPurchases[value,1] <- sub(pattern = "^u.*", replacement = "Unilever", x = prodPurchases[value,1])}

test <- grep(pattern = "^v.*", prodPurchases$company)
for(value in test) {prodPurchases[value,1] <- sub(pattern = "^v.*", replacement = "Van Houten", x = prodPurchases[value,1])}
```
```{r echo=FALSE}
prodPurchases$company
```

Just for good measure, let's make sure we don't have any left over.
```{r}
prodPurchases %>% count(company)
```


Next, the product codes and product numbers need to be separated to make viewing and working with this data easier.
```{r}
prodPurchases <- separate(prodPurchases, Product.code...number, c("product_code", "product_number"), sep = "-")
```
```{r echo=FALSE}
kable(head(select(prodPurchases, product_code, product_number))) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```


## What's missing?
Now I will add a column for product_category that's easier to read.
```{r}
prodPurchases$product_category <- sub(pattern = "^p$", replacement = "Smartphone", x = sub("^x$", "Laptop", sub("^v$", "TV", sub("^q$", "Tablet", prodPurchases$product_code))))
```
```{r echo=FALSE}
kable(head(select(prodPurchases, product_code, product_category))) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```

Now I'll add a column for geocoding called full_address
```{r}
prodPurchases <- prodPurchases %>% 
  mutate(full_address = paste(address, city, country, sep = ","))
```
```{r echo=FALSE}
kable(head(select(prodPurchases, address, city, country, full_address))) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```

I will also add 8 more columns, all binary representing what companies or products are referenced in that record.
```{r}
prodPurchases <- mutate(prodPurchases, company_philips = ifelse(company == "Philips", 1, 0))
prodPurchases <- mutate(prodPurchases, company_akzo = ifelse(company == "Akzo", 1, 0))
prodPurchases <- mutate(prodPurchases, company_van_houten = ifelse(company == "Van Houten", 1, 0))
prodPurchases <- mutate(prodPurchases, company_unilever = ifelse(company == "Unilever", 1, 0))
prodPurchases <- mutate(prodPurchases, product_smartphone = ifelse(product_category == "Smartphone", 1, 0))
prodPurchases <- mutate(prodPurchases, product_tv = ifelse(product_category == "TV", 1, 0))
prodPurchases <- mutate(prodPurchases, product_laptop = ifelse(product_category == "Laptop", 1, 0))
prodPurchases <- mutate(prodPurchases, product_tablet = ifelse(product_category == "Tablet", 1, 0))
```
```{r echo = FALSE}
kable(prodPurchases %>% 
  select(company, company_philips, company_akzo, company_unilever, company_van_houten) %>%
    head()) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```
```{r echo=FALSE}
kable(prodPurchases %>% 
  select(product_category, product_smartphone, product_tv, product_laptop, product_tablet) %>%
    head()) %>% 
  kable_styling(bootstrap_options = c("striped","hover", "condensed"))
```

## Finish up!
Finally, write a new .csv file for review
```{r}
write.csv(prodPurchases, "refine_clean.csv")
```






